# Week 4

---
title: "Gender Inequality Index"
author: "Burhan"
date: "2023-10-31"
output: html_document
---

# Load Packages

```{r}
library(tidyverse)
library(sf)
library(here)
library(janitor)
library(countrycode)
```

## Loading Spatial File (as Shape File .shp/.geojson) & Data File (as Comma Separated Value File .csv)

```{r}

world_map <- st_read (here::here("Data/World_Countries_Generalized.geojson"))


HDI_csv <- read_csv(here::here("Data", "HDR21-22_Composite_indices_complete_time_series.csv"),
                       locale = locale(encoding = "latin1"),
                       na = "", skip=0)
```

## Checking the data type in both of my files (i.e. rows, columns, type of data, etc)

```{r}
head(world_map)
head(HDI_csv)
```

## Plotting the shape file to check the geometry

```{r}
world_map %>% 
  st_geometry() %>%
  plot()
```

## Cleaning both the files

```{r}
HDI_myclean_csv <- clean_names(HDI_csv)
world_map <- clean_names(world_map)
```

## Adding the three code country code data to our geometry file using countrycode

```{r}
world_map_iso <- world_map %>% 
  mutate(iso_country_code = countrycode(country, origin = 'country.name', destination = 'iso3c'))
head(world_map_iso)
```

## Left join the two data files using the iso codes available in both data sets

```{r}
final_join <- world_map_iso %>% 
  left_join(., 
            HDI_myclean_csv,
            by = c("iso_country_code" = "iso3")) #use shape file column = csv column

```

## Now make a new column in the data set by obtaining the difference in Gender Inequality Index from 2019 to 2010

```{r}
final_join_diff <- final_join %>% 
  mutate(difference=gii_2019-gii_2010)
```

## Plot the data

```{r}
plot1 <-
  ggplot(data = final_join_diff) +
  geom_sf(aes(fill = difference)) +
  scale_fill_viridis_c(name = "Gender Inequality Index Diff") +
  labs(title = "Gender Inequality b/w 2010 and 2019") +
  theme_minimal() +
  theme(legend.position = "right")

plot1
```

## Save Plot as image

image modification before saving

```{r}
# change the theme to save
save_map <- plot1 + 
  theme(
    text = element_text(
      size = 20,
    ),
    plot.title = element_text(
      size = 20,
    ),
    legend.title = element_text(size = 20),
  ) 
  
ggsave(
  "Images/Result_GII.png",
  plot = save_map,
  height = 35,
  width = 35)

```
