# Practical 2

---
title: "London_Boroughs_Work"
author: "Burhan"
date: "2023-10-11"
output: html_document
---

# Loading Packages

```{r}
library(sf)
library(tmap) 
library(tmaptools)
library(RSQLite)
library(tidyverse)
library(dplyr)
library(janitor)
library(rJava)
library(OpenStreetMap)
library(NLP)
```

## Loading Spatial File (as Shape File .shp) & Data File (as Comma Separated Value File .csv)

```{r}
shape_lon <- st_read("C:/Users/Burhan/Documents/GIS_TERM1_UCL/R_Directory/R_Tasks_UCL/Data/Local_Authority_Districts_(December_2015)_Boundaries/Local_Authority_Districts_(December_2015)_Boundaries/Local_Authority_Districts_(December_2015)_Boundaries.shp")

mycsv_lon <- read_csv("C:/Users/Burhan/Documents/GIS_TERM1_UCL/R_Directory/R_Tasks_UCL/Data/ward-profiles-excel-version.csv",locale =locale(encoding="latin1"),na="n/a")
```

## Getting a glimpse of the data type in both of my files (i.e. rows, columns, type of data, etc)

```{r}
# glimpse(mycsv_lon)
# # it only gives you a view of data and cannot create a data set out of it
# glimpse(shape_lon)
```

## Checking the summarise function of dplyr vs glimpse used above

```{r}
# class(mycsv_lon)
# X <- mycsv_lon %>% 
#   summarise_all(class) %>%
#   pivot_longer(everything(), 
#                names_to="All_variables", 
#                values_to="Variable_class")
# X
```

## Checking the first 10 column names using head & colnames functions

```{r}
# mycsv_lon %>% 
#   colnames() %>% 
#   head(10)
```

## Using Janitor for cleaning the data it replaces all spaces with "\_" and turns all data in shortcase

```{r}
mycsv_clean <- clean_names(mycsv_lon)
```

## Checking what janitor did to thr column names

```{r}
# mycsv_clean %>% 
#   colnames() %>% 
#   head(10)
```

## Plotting the shape file to check the geometry

```{r}
# shape_lon %>% 
#   st_geometry() %>%
#   plot(.)
```

## Adding the attributes of Data File to Spatial File by comparing the 'lad15cd' column in spatial file with the 'new_code' column in data file

```{r}
shape_lon <- shape_lon%>%
  merge(.,
        mycsv_clean,
        by.x="lad15cd", 
        by.y="new_code")
```

## Plotting the map by using the "Paid Employee" attribute

```{r}
# tmap_mode("plot")
# shape_lon %>% 
#   qtm(., fill = "population_2015")
```

## Plotting with ggplot2

```{r}
ggplot(shape_lon) +
  geom_sf(aes(fill = children_aged_0_15_2015)) +
  scale_fill_viridis_c(name = "children_2015") +
  labs(title = "London Boroughs Map", fill = "Population") +
  theme_minimal() +
  theme(legend.position = "right")
```

## Now get only London Boroughs from mycsv_clean

```{r}
LondonBoroughs<- mycsv_clean %>% 
  filter(str_detect(`new_code`, "^E09"))
```

## checking whether the filter of E09 worked

```{r}
# LondonBoroughs %>% 
#   dplyr::select(`ward_name`) %>%
#   print()
```

## As City of London is coming twice so only keeping distint values

```{r}
LondonBoroughs<-LondonBoroughs %>%
  distinct()
```

## Now manually selecting the columns we require from the big spreadsheet

```{r}
LondonBoroughs_dplyrcols<-LondonBoroughs %>%
  dplyr::select(c(1,19,20,21))
```

## Trying the same with contains dplyr tool

```{r}
LondonBoroughs_contains<-LondonBoroughs %>% 
  dplyr::select(contains("ward_name"),
                contains("expectancy"), 
         contains("obese_2011_12_to_2013_14")) 
```

# Making a new data set with two new columns which we obtain from the LondonBorough data set

```{r}
Life_expectancy <- LondonBoroughs %>%
  mutate(AVERAGELIFEEXPECTANCY=(female_life_expectancy_2009_13+male_life_expectancy_2009_13)/2)%>%
  mutate(NORMALISEDLIFEEXPECTANCY= AVERAGELIFEEXPECTANCY /
           mean(AVERAGELIFEEXPECTANCY))
#new column with average of male and female life expectancy
```

## Checking whether the columns are added in the new dataset.

```{r}
# glimpse(Life_expectancy)
```

## Now only selecting the columns which we required.

```{r}
Life_expectancy %>% 
dplyr::select(new_code, ward_name,AVERAGELIFEEXPECTANCY,NORMALISEDLIFEEXPECTANCY)%>%
  arrange(desc(NORMALISEDLIFEEXPECTANCY))
```

## We can also check the top 5 rows of the data using

```{r}
#top of data
# slice_head(Life_expectancy, n=5)
```

## We can also check the bottom 5 rows of the data using

```{r}
#top of data
# slice_tail(Life_expectancy, n=5)
```

## Now we use dplyr to check the data above and below UK Average of 81.16 and then assign boroughs as per that criteria "above average" or "below average"

```{r}
Life_expectancy2 <- Life_expectancy %>%
  mutate(UKcompare = case_when(AVERAGELIFEEXPECTANCY>81.16 ~ "above UK average",
                               TRUE ~ "below UK average"))
# Life_expectancy2
```

# Now we want to summarize this by creating a new data set and grouping it by "above or below average criteria and finally summarizing it by obtaining max-min difference, count and average.

```{r}
Life_expectancy2_group <- Life_expectancy2 %>%
  mutate(UKdiff = AVERAGELIFEEXPECTANCY-81.16) %>%
  group_by(UKcompare)%>%
  summarise(range=max(UKdiff)-min(UKdiff), count=n(), Average=mean(UKdiff))

# Life_expectancy2_group
```

## Further analysis to obtain diffrence by years (IMPORTANT)

```{r}
Life_expectancy3 <- Life_expectancy %>%
  mutate(UKdiff = AVERAGELIFEEXPECTANCY-81.16)%>%
  mutate(across(where(is.numeric), round, 3))%>%
  mutate(across(UKdiff, round, 0))%>%
  mutate(UKcompare = case_when(AVERAGELIFEEXPECTANCY >= 81 ~ 
                                 str_c("equal or above UK average by",
                                       UKdiff, 
                                       "years", 
                                       sep=" "), 
                               TRUE ~ str_c("below UK average by",
                                            UKdiff,
                                            "years",
                                            sep=" ")))%>%
  group_by(UKcompare)%>%
  summarise(count=n())
```

```{r}
Life_expectancy4 <- Life_expectancy %>%
  mutate(UKdiff = AVERAGELIFEEXPECTANCY-81.16)%>%
  mutate(across(is.numeric, round, 3))%>%
  mutate(across(UKdiff, round, 0))
```

## Now we will go for plotting

```{r}
plot(LondonBoroughs$male_life_expectancy_2009_13,
     LondonBoroughs$percent_children_in_reception_year_who_are_obese_2011_12_to_2013_14)
```

```{r}
# install.packages("maptools")
# install.packages(c("classInt", "tmap"))
# install.packages(c("RColorBrewer", "sp", "rgeos", 
#                    "tmaptools", "sf", "downloader", "rgdal", 
#                    "geojsonio"))
# install.packages("RgoogleMaps")
# install.packages("Openstreetmaps")
```

```{r}
#library(maptools)
library(RColorBrewer)
library(classInt)
library(sp)
#library(rgeos)
library(sf)
#library(rgdal)
library(geojsonio)
library(ggplot2)
library(viridis)
library(RgoogleMaps)
```

## Saving the shape file in LondonMap and removing duplicates

```{r}
LondonMap <- shape_lon %>%
  filter(str_detect(`lad15cd`, "^E09")) %>%
  distinct()
```

```{r}
qtm(LondonMap)
```

```{r}
LondonMap <- LondonMap%>%
  left_join(., 
            Life_expectancy4,
            by = c("lad15cd" = "new_code"))
```

```{r}
# palette_ucl <- c("#500778", "#e0ecf4","#FFFFFF","#8856a7")
```

```{r}
ggplot(LondonMap) +
  geom_sf(aes(fill =rate_of_job_seekers_allowance_jsa_claimants_2015.x)) +
  scale_fill_viridis_c(name = "Rate of Claims") +
  labs(title = "London Map- JSA -2015") +
  theme_minimal() +
  theme(legend.position = "right")
```
## Let us try same with tmaptools qtm tool

```{r}
library(tmap)
library(tmaptools)
library(sf)
tmap_mode("plot")
qtm(LondonMap, 
    fill = "rate_of_job_seekers_allowance_jsa_claimants_2015.x")
```
## Add a base map now & save it in tmaplondon
```{r}
tmaplondon <- LondonMap %>%
  st_bbox(.) %>% 
  tmaptools::read_osm(., type = "osm", zoom = NULL)
```

## Use a basemap with our plot 

```{r}
tmap_mode("plot")
tm_shape(tmaplondon)+
tm_rgb()+
tm_shape(LondonMap) + 
tm_polygons("rate_of_job_seekers_allowance_jsa_claimants_2015.x", 
        style="jenks",
        palette="YlOrBr",
        midpoint=NA,
        title="Rate per 1,000 people",
        alpha = 0.5) + 
  tm_compass(position = c("left", "bottom"),type = "arrow") + 
  tm_scale_bar(position = c("left", "bottom")) +
  tm_layout(title = "Job seekers' Allowance Claimants", legend.position = c("right", "bottom"))
```
## Making a new map with the help of inner join with our analysed data set i.e. Life_expectancy4 & check which columns got added to our original file shape_lon

```{r}
Life_expectancy4map <- shape_lon %>%
    inner_join(., 
            Life_expectancy4,
            by = c("lad15cd"="new_code"))%>%  #first keep (shape file column = excel sheet column i.e. Life_expectancy4 here)
    distinct(.,lad15cd, 
           .keep_all = TRUE)
```

## Plotting final result

```{r}
tmap_mode("plot")
      tm_shape(tmaplondon)+
      tm_rgb()+
      tm_shape(Life_expectancy4map) +
      tm_polygons("UKdiff", 
        style="pretty",
        palette="Greens",
        midpoint=NA,
        title="Number of years",
        alpha = 0.5) + 
       tm_compass(position = c("left", "bottom"),type = "arrow") + 
       tm_scale_bar(position = c("left", "bottom")) +
       tm_layout(title = "Difference in life expectancy",legend.position = c("right", "bottom"))

```
# Same with ggplot 
```{r}
# l <-ggplot(LondonMap) +
#   geom_sf(aes(fill =UKdiff)) +
#   scale_fill_viridis_c(name = "UK Difference(Borough vs UK)") +
#   labs(title = "Difference in Life Expectancy") +
#   theme_minimal() +
#   theme(legend.position = "bottom")

l
```


